### 2022-09-14 README.Rmd
Packaged checked 1 warning 2 notes

---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# brazildatamortality

<!-- badges: start -->
<!-- badges: end -->

This package contains both the data and the figures presented in the paper 
"Natural  Hazards Fatalities in Brazail, 1979–2019". Below is the abstract of the
aforementioned paper.

The impact of natural hazards on nations and societies is a global challenge and
concern. Studies worldwide have been conducted within and among countries, to 
examine the spatial distribution and temporal evolution of fatalities and their 
consequences in societies. In Brazil, no studies have comprehensively identified 
fatalities associated with all natural hazards and their singularities by 
decade, region, sex, age, and other victim characteristics. This study develops 
a deep analysis on the Brazilian Data Mortality of the Brazilian Ministry of 
Health, from 1979 to 2019, identifying the natural hazards that kill the 
greatest number of people in Brazil and its surrounding particularities. 
Lightning is the deadliest natural hazard in Brazil during this period, with a 
gradual decrease in the number of fatal victims. Hydrogeological fatalities 
increases from 2000, and the most fatalities develop from 2010 to 2019. Despite 
Brazil being a tropical country affected by severe droughts, extreme heat had 
the lowest number of fatalities, almost irrelevant when compared with that of 
other natural hazards. The period from December to March is with the higher 
number of fatalities, and the Southeast region is the most populous regions were 
the most are fatally affected. The number of male victims is double that of 
female victims, of all ages, and unmarried victims died the most. Thus, it is 
fundamental to recognize and make public the knowledge of different natural 
hazards’ impacts on communities and societies, namely people and their 
livelihoods, to evaluate challenges and recognize opportunities to reduce 
natural hazards’ impacts on Brazil.

## Installation

You can install the development version of brazildatamortality from 
[GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("coutogabriela/brazildatamortality")
```

## Example

This basic example shows the deaths in Brazil by cause and year:

```{r example}

library(brazildatamortality)
library(dplyr)
library(ggplot2)

get_data() %>%
    ggplot() +
    geom_bar(aes(x = death_year, fill = cause)) +
    xlab("Year") +
    ylab("Number of fatalities") +
    labs(fill = "Fatality cause") +
    ggtitle("Deaths by natural hazard in Brazil")

```

This package contains two sets of data corresponding to fatalities by natural
hazards (one from 1979 to 1995 and the othre from 1996 to 2019), two datasets corresponding to Brazil's administrative divisions (state and town) and a last
dataset corresponding to a hexagonal approximation of the state division.

### 2022-09-27 README.Rmd
Typos fixed.

@@
 This package contains both the data and the figures presented in the paper 
-"Natural  Hazards Fatalities in Brazail, 1979–2019". Below is the abstract of the
+"Natural  Hazards Fatalities in Brazil, 1979–2019". Below is the abstract of the
 aforementioned paper.
@@
 
-This package contains two sets of data corresponding to fatalities by natural
-hazards (one from 1979 to 1995 and the othre from 1996 to 2019), two datasets corresponding to Brazil's administrative divisions (state and town) and a last
-dataset corresponding to a hexagonal approximation of the state division.
+This package contains:
+
+- two datasets corresponding to fatalities by natural hazards (one from 1979 to 1995 
+and the other from 1996 to 2019). 
+- two datasets corresponding to Brazil's administrative divisions (state and town) 
+- one dataset corresponding to a hexagonal approximation of the Brazilian state 
+division.
